---
title: "dwplot: Plot Comparable Estimates"
author: "Frederick Solt and Yue Hu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: vignette_dwplot.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{ASCII}
---

The `dwplot` package provides a way to create highly customizable plot for presenting and comparing statistics. Users can use it to plot coefficients or other estimates (e.g., predicted probabilities) within a model or compare them across different models. The estimates are presented as dots with confidential interval whiskers, and grouped based on variables.

Users can easily custmize the number of models, the exact variables present, and the inverval between the compared estimates based on their interests and presenting demands. Moreover, by outputting `ggplot` objects, `dwplot` allows users to easily further manage their graphs.

This vignette purposes to illustrate how users can customize the presenting data frame, and produce decent plots.


## Create a Data Frame of Estimations

```{r message = F}
library(dplyr)
library(broom)
library(ggplot2)
```

### Standardized Results
@Gelman2008 suggest that 


## Decently Use `dwplot`

### Plot Coeffients


### Plot Predicted Probabilities


### Plot Models with Different Variables

```{r}
# The "Secret Weapon" (plot one coefficient for many models)
data(diamonds)

by_clarity <- diamonds %>% group_by(clarity) %>% 
    do(tidy(lm(price ~ carat + cut + color, data = .))) %>% rename(model=clarity) 

carat_results <- by_clarity %>% filter(term=="carat") %>% select(-term) %>% 
    rename(term=model) 

#dwplot(carat_results) + 
#    xlab("Estimated Coefficient (Dollars)") + ylab("Diamond Clarity") + 
#    ggtitle("Estimated Coefficients for Diamond Size Across Clarity Grades") +
#    theme(plot.title = element_text(face="bold"), 
#          legend.position = "none")
```
